\documentclass[12pt, fullpage, letterpaper]{article}
\usepackage{fancyheadings, graphicx}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nsf}
\renewcommand{\headrulewidth}{0pt}
\newcommand{\doi}[2]{#2}
\pagestyle{empty}

\begin{document}

\subsection*{David W. Hogg}
\noindent http://cosmo.nyu.edu/hogg/

\subsubsection*{Professional preparation}
\begin{list}{}{\hogglist}
\item
SB (Physics) 1992, Massachusetts Institute of Technology
\item
PhD, Physics, 1998, California Institute of Technology
\item
Long-term member, 1997--2001, Institute for Advanced Study
\end{list}

\subsubsection*{Appointments}
\begin{list}{}{\hogglist}
\item
\deemph{current:} Professor of Physics and Data Science, New York University, 2014--.
\item
\deemph{current:} Group Leader, Astronomical Data Group, Center for Computational Astrophysics,
Flatiron Institute, New York, 2017--.
\item
\deemph{current:} Adjunct Senior Staff Scientist, Max-Planck-Institut f\"ur Astronomie,
Heidelberg, Germany, 2012--.
\item
Consultant, Flatiron Institute, New York, 2016--2017.
\item
Consultant, Simons Center for Data Analysis,
Simons Foundation, New York, 2015--2016.
\item
Associate Professor of Physics with tenure, New York University, 2007--2014.
\item
Visiting Scientist, Max-Planck-Institut f\"ur Astronomie,
Heidelberg, Germany, 2006--2011.
\item
Visiting Professor, Department of Astronomy and Astrophysics, Columbia
University, 2008
\item
Assistant Professor of Physics, New York University, 2001--2007.
\end{list}

\subsubsection*{Products---five relevant} % 5
\begin{list}{}{\hogglist}
\item
Hogg,~D.~W., Eisenstein,~D.~J., Blanton,~M.~R., Bahcall,~N.~A., Brinkmann,~J., Gunn,~J.~E., \& Schneider,~D.~P., 2005,
\doi{10.1086/429084}{Cosmic homogeneity demonstrated with luminous red galaxies},
\textit{Astrophys.\,J.}\ \textbf{624} 54--58.
\item\label{pub:Eisenstein2005}
Eisenstein,~D.~J., Zehavi,~I., Hogg,~D.~W., \etal, 2005,
\doi{10.1086/466512}{Detection of the baryon acoustic peak in the large-scale correlation function of \project{Sloan Digital Sky Survey} Luminous Red Galaxies},
\textit{Astrophys.\,J.}\ \textbf{633} 560--574.
\item\label{pub:Masjedi2006}
Masjedi,~M., Hogg,~D.~W., \etal, 2006,
\doi{10.1086/503536}{Very small-scale clustering and merger rate of luminous red galaxies},
\textit{Astrophys.\,J.}\ \textbf{644} 54--60.
\item
Padmanabhan,~N., \etal, 2008,
\doi{10.1086/524677}{An improved photometric calibration of the \project{Sloan Digital Sky Survey} imaging data},
\textit{Astrophys.\,J.}\ \textbf{674} 1217--1233.
%\item
%More,~S., Bovy,~J., \& Hogg,~D.~W., 2009,
%\doi{10.1088/0004-637X/696/2/1727}{Cosmic transparency:\ A test with the baryon acoustic feature and type Ia supernovae},
%\textit{Astrophys.\,J.}\ \textbf{696} 1727--1732.
\item
Hahn,~C., Vakili,~M., Walsh,~K., Hearin,~A.~P., Hogg,~D.~W., \& Campbell,~D., 2017,
\doi{10.1093/mnras/stx894}{Approximate Bayesian computation in large-scale structure: constraining the galaxy-halo connection},
\textit{Mon.\,Not.\,R.\,Astr.\,Soc.}\ \textbf{469} 2791--2805.
\end{list}

\subsubsection*{Products---five other significant} % 5
\begin{list}{}{\hogglist}
\item
Willman,~B., Blanton,~M.~R., West,~A.~A., Dalcanton,~J.~J, Hogg,~D.~W., Schneider,~D.~P., Wherry,~N., Yanny,~B., \& Brinkmann,~J., 2005,
\doi{10.1086/430214}{A new Milky Way companion:\ Unusual globular cluster or extreme dwarf satellite?},
\textit{Astron.\,J.}\ \textbf{129} 2692--2700.
\item
Lang,~D., Hogg,~D.~W., Mierle,~K., Blanton,~M., \& Roweis,~S., 2010,
\doi{10.1088/0004-6256/139/5/1782}{\project{Astrometry.net}:\ Blind astrometric calibration of arbitrary astronomical images},
\textit{Astron.\,J.}\ \textbf{139} 1782--1800.
\item
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011,
\doi{10.1214/10-AOAS439}{Extreme deconvolution:\ Inferring complete distribution functions from noisy, heterogeneous and incomplete observations},
\textit{Ann.\,Appl.\,Stat.}\ \textbf{5} 1657--1677.
\item
Ness,~M., Hogg,~D.~W., Rix,~H.-W., Ho,~A.~Y.~Q., \& Zasowski,~G., 2015,
\doi{10.1088/0004-637X/808/1/16}{\project{The~Cannon}: A data-driven
approach to stellar label determination},
\textit{Astrophys.\,J.}\ \textbf{808} 16.
\item
Bonaca,~A., Hogg,~D.~W., Price-Whelan,~A.~M., \& Conroy,~C., 2019,
\doi{10.3847/1538-4357/ab2873}{The Spur and the Gap in GD-1:\ Dynamical evidence for a dark substructure in the Milky Way halo},
\textit{Astrophys.\,J.}\ \textbf{880} 38.
\end{list}

\subsubsection*{Synergistic activities} % up to 5
\begin{list}{}{\hogglist}
\item
Helped the astronomical community through advisory committees:
\project{Spitzer} Science Center Oversight Committee (to advise the \project{Spitzer Space Telescope} project) 2008--present.
\project{\NASA\ Extragalactic Database} Users Committee 2006--2013.
Astronomy and Astrophysics Advisory Committee (established by the US
Congress to advise \NSF, \NASA, and \DOE\ on areas of overlap) 2014--2017.
\item
Photometrically calibrated the \project{Sloan Digital Sky Survey} imaging data; this
calibration has been used in all data releases since DR8.
Involved in the every phase of the
\project{SDSS-IV} project. Member of the
\project{SDSS-IV} Collaboration Council,
2013--present. Currently helping with the design of \project{SDSS-V}.
%% \item
%% Served in administrative roles at NYU, including
%% Director of Undergraduate Studies, Department of Physics, New York University, 2008--2015; and
%% Executive Director of the \textsc{usd}12.6M \project{Moore--Sloan Data Science Environment at \textsc{nyu}}, 2013--2015.
\item
Supervised the production of multiple multi-user open-source software projects,
including \project{Astrometry.net} for blind image calibration (thousands of users),
\project{Daft} for drawing graphical models, and the \project{emcee}
MCMC sampler (more than 3500 citations in 6 years).
\item
Wrote and maintain pedagogical documents about a variety of topics including
cosmological distance measures (arXiv:astro-ph/9905116),
probability (arXiv:1205.4446),
fitting models (arXiv:1008.4686), and
MCMC (arxiv:1710.06068).
\item
Designed, organized, and promoted hack events for education and scientific research:
Co-founded and co-organized the \project{AstroHackWeek} series.
Founded and co-organized the \project{Hack Together Day} at the Winter AAS meetings.
Founded, designed, and co-organized the \project{Gaia Sprints}.
Co-authored a paper on some of these activities (arXiv:1711.00028).
\end{list}

\end{document}
