\documentclass[12pt, fullpage, letterpaper]{article}
\usepackage{fancyheadings, graphicx}
\setlength{\headheight}{0ex}
\setlength{\headsep}{0ex}
\input{hogg_nsf}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{empty}

\begin{document}

\subsection*{Data management plan}

This project will not produce new data \foreign{per se}, but it does
produce methods, code, statistical results, and value-added
calibration models for some large cosmology projects.  All of the
code, documentation, data, and all written documents relating to the
project will be maintained in publicly visible version-control
repositories on the GitHub or GitLab platforms (as is this
proposal). These repositories can be cloned and reproduced by any user
in the world.

The code and documents will be licensed for re-use by others under permissive
licenses (MIT for code, CC-by for documents). This permits re-use, reproducibility,
and re-publication by others.
We will ensure that code and data are snapshotted and archived with Zenodo for
preservation and the creation of cite-able DOIs.
All scientific  papers will be submitted to the arXiv and also \AAS-family
journals for long-term preservation, cite-ability, and publication.
All pedagogical writing will be submitted to the arXiv, and, where appropriate,
refereed journals.

The PI has a strong tradition in his group for open science, performing
all data analysis and writing in the open, with a strong preference for working
in open and public data sets. He also supports maintainance and development of
open-source software systems of wide use in the astronomical community.

\end{document}
